---
title: 'MTH-2520 Homework 8: Group Manipulation, User Input and Dating'
author: Brady Lamson
date: '2021-03-25'
slug: mth-2520-homework-8-group-manipulation-user-input-and-dating
categories: []
tags: [Homework]
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="")

library(rmarkdown, tinytex)
library(markdown)
library(dplyr)
library(stringr)
library(ggplot2)
library(reshape)
library(lubridate)
```


## Problem 1
The \textbf{geometric mean} of a vector of numbers $c(a1, a2, a3, ..., an)$ is given by 

\[\text{gmean} = (a_1 \cdot a_2 \cdot a_3 \cdot ... \cdot a_n)^{\frac{1}{n}}\]

### Problem 1a
Write a user-defined function "gmean" that computes the geometric mean of a vector of numbers.

```{r}
gmean <- function(vector) {
  vec_len <- length(vector)
  
  ret_val <- prod(vector)^(1 / vec_len)
  
  ret_val
}
```


### Problem 1b
Test your function on the following vectors
```{r}
pr1a <- c(4,9)
gmean(pr1a)
```

```{r}
pr1b <- c(2,8)
gmean(pr1b)
```

```{r}
pr1c <- c(5,9)
gmean(pr1c)
```

```{r}
pr1d <- c(4,1,1/32)
gmean(pr1d)
```

```{r}
pr1e <- c(8,10,51.2)
gmean(pr1e)
```

```{r}
pr1f <- c(1,3,9,27,81)
gmean(pr1f)
```


\pagebreak


## Problem 2

### Problem 2a
Construct a random 5x6 matrix with $n=30$ elements:

```{r}
A <- matrix(sample(30, replace = TRUE), nrow = 5)
```

### Problem 2b
Use the apply() to find the geometric mean of each column

```{r}
cat("The mean of each column is [", round(apply(A, MARGIN = 2, FUN = gmean), digits = 3), "] \n")
```


\pagebreak


## Problem 3 

### Problem 3a
Construct a list containing the following three matrices as elements:

```{r}
a3 <- matrix(1:9, nrow = 3)
b3 <- matrix(4:15, nrow = 4)
c3 <- matrix(8:13, nrow = 3)

pr3_list <- list(a3, b3, c3)
```

```{r}
pr3_list
```


### Problem 3b
Use lapply() to extract the 2nd column of each matrix from the list

```{r}
lapply(pr3_list, FUN = "[", , 2) #Essentially: lapply(data, FUN = "[", rows=all, columns=2)
```
To briefly explain what's happening here, the function I'm using within lapply is "[". This function takes in two arguments. It takes in rows, and then columns. So I left an empty space within the commas to include all rows and then specified the second column. This feels a little gross and I can't help but feel as if I'm missing a far more intuitive solution. 


### Problem 3c
Use lapply() to extract the 1st row of each matrix from the list

```{r}
lapply(pr3_list, FUN = "[", 1, ) #Essentially: lapply(data, FUN = "[", rows=1, columns=all)
```


### Problem 3d
Use sapply() to find the row sums of each matrix in the list.

```{r}
sapply(pr3_list, rowSums)
```


### Problem 3e
Use sapply() to find the geometric mean of each matrix in the list.

```{r}
cat("The geometric mean of each matrix is [", 
    round(sapply(pr3_list, gmean), digits=3), 
    "] \n")
```


\pagebreak


## Problem 4
From the Frogger data frame created in the Week 9 lab, compute the mean score and the geometric mean score. 

```{r}
# Create the frogger data frame
frogger <- data.frame(
	Player=rep(c("Alice", "Bob", "Connor"), times=c(2,5,3)), 
	Score=round(rlnorm(n=10,10), -1))
```

```{r}
# Compute mean score
mean(frogger$Score)
```

```{r}
# Compute geometric mean score
cat("The geometric mean of the frogger scores is",
  gmean(frogger$Score),
  "\n")
```


\pagebreak


## Problem 5
Import the built-in dataset mtcars.

```{r}
data("mtcars")
```

### Problem 5a
Check the data types and names of the columns. Check for NAs.

```{r}
colnames(mtcars)
```

```{r}
sapply(mtcars, typeof)
```

```{r}
sapply(mtcars, function(x) sum(is.na(x)))
```


### Problem 5b
Split the mpg data based on the number of cylinders.

```{r}
# with(mtcars, split(mpg, cyl))
split(mtcars$mpg, mtcars$cyl) # I much prefer the syntax on this one.
```


### Problem 5c
Use tapply() to compute the mpg based on the number of cylinders.

```{r}
tapply(mtcars$mpg, mtcars$cyl, mean)
# with(mtcars, tapply(mpg, cyl, mean)) Here just to show alternate solution
```


### Problem 5d
Use tapply() to compute the mean horse power hp basedd on the number of cylinders.

```{r}
tapply(mtcars$hp, mtcars$cyl, mean)
```


### Problem 5e
Use apply() combined with tapply() to compute the mean of each column based on the number of cylinders

```{r}
apply(mtcars, MARGIN = 2, function(x) tapply(x, mtcars$cyl, mean))
```


\pagebreak


## Problem 6
Write your own Mad Lib that takes in at least 5 input words or numbers, and is at least a few sentences long. (Use a fixed input vector in the .rmd printout that you turn in) You can use a well-known poem, passage, or speech, or make up your own story.


```{r}
# Inputs: noun, color, noun2, verb, color

x <- list(noun = "velociraptor", color = "lavender", pluralnoun = "puppies", verb = "vibin", verb = "nuzzle", color = "vanilla", verb = "cuddle")

cat("Sometimes that", x[[1]], "looks right at ya. Right into your eyes. And the thing about a", x[[1]], "is he's got lifeless eyes,", x[[2]], "eyes. Like a", x[[3]], "eyes. \n", "When he comes at ya, he doesn't even seem to be", x[[4]], "till he", x[[5]], " ya and those", x[[2]], "eyes roll over white and then ah then you hear that terrible high-pitched screamin'. The ocean turns", x[[6]], "and despite all your poundin' and your hollerin' those", x[[1]], "come in and they", x[[7]], "you to pieces. \n")
```

\pagebreak


## Problem 7
Use the appropriate libridate function to parse each of the following dates. Also, find the weekday for each date. 

### Problem 7a

```{r}
d1 <- "February 2, 2021"

d1_x <- mdy(d1)
d1_x

wday(d1_x, label=TRUE)
```


### Problem 7b

```{r}
d2 <- "4/1/2025" #American date. Goal is April 1st

d2_x <- mdy(d2)
d2_x

wday(d2_x, label=TRUE)
month(d2_x, label=TRUE)
```


### Problem 7c

```{r}
d3 <- "4/1/2025" #European date. Goal is Jan 4th

d3_x <- dmy(d2)
d3_x

wday(d3_x, label=TRUE)
month(d3_x, label=TRUE)
```


### Problem 7d

```{r}
d4 <- "2019-Jun-30"

d4_x <- ymd(d4)
d4_x

wday(d4_x, label = TRUE)
month(d4_x, label = TRUE)
```


### Problem 7e

```{r}
d5 <- "11 Nov 2011"

d5_x <- dmy(d5)
d5_x

wday(d5_x, label = TRUE)
month(d5_x, label = TRUE)
```


\pagebreak


## Problem 8


### Problem 8a
Compute the time since the first moon landing. Covert this output to years. 

```{r}
age <- today() - mdy("7/20/1969")
as.duration(age)
```


### Problem b
Create the following data frame. 

```{r}
# We're given an example data frame that I need to recreate before I can get started. 

Apollo_Mission <- c(11, 12, 14, 15, 16, 17)

Launch_Date <- c("16 July 1969", "14 Nov 1969", "31 Jan 1971", "26 July 1971", "16 Apr 1972", "07 Dec 1972")

Launch_Time <- c("13:32", "16:22", "21:03", "13:34", "17:54", "05:33")

Moon_L_Date <- c("7/20/1969", "11/19/1969", "2/5/1971", "7/30/1972", "4/21/1972", "12/11/1972")

Moon_L_Time <- c("20:17:40", "06:54:35", "09:18:11", "22:16:29", "02:23:35", "19:54:57")

Splashdown_L_Date <- c("Jul 24, 1969", "Nov 24, 1969", "Feb 9, 1971", "Aug 7, 1971", "Apr 29, 1972", "Dec 19, 1972")

Splashdown_L_Time <- c("16:50:35", "20:58:24", "21:05:00", "21:45:53", "19:45:05", "19:24:59")

apollo <- dplyr::tibble(Apollo_Mission, Launch_Date, Launch_Time, Moon_L_Date, Moon_L_Time, Splashdown_L_Date, Splashdown_L_Time)
apollo
```


### Problem c
Convert all the information to standard lubridate dates and times
```{r}
# unite columns to get date/hour/minute/second format-------------------
apollo <- tidyr::unite(apollo, "Moon_Landing", Moon_L_Date:Moon_L_Time)
apollo <- tidyr::unite(apollo, "Launch", Launch_Date:Launch_Time)
apollo <- tidyr::unite(apollo, "Splashdown", Splashdown_L_Date:Splashdown_L_Time)

# Convert times to something usable-------------------------------------
apollo$Launch <- dmy_hm(apollo$Launch)
apollo$Moon_Landing <- mdy_hms(apollo$Moon_Landing)
apollo$Splashdown <- mdy_hms(apollo$Splashdown)
```

```{r}
apollo
```


### Problem d
Compute the duration of each mission in days.
(Splashdown time - Launch time)
```{r}
mission_duration <- apollo$Splashdown - apollo$Launch
mission_table <- dplyr::tibble(apollo$Apollo_Mission, mission_duration)
```

```{r}
mission_table
```


### Problem 8e
Compute the average duration of a mission

```{r}
cat("The average mission has a time difference of", mean(mission_table$mission_duration), "days \n")
```

