---
title: 'MTH-2520 Homework 10: Graphing with ggplot2'
author: Brady Lamson
date: '2021-04-13'
slug: mth-2520-homework-10-graphing-with-ggplot2
categories: []
tags: [Homework]
---

```{r setup, include = FALSE}
#knitr::opts_chunk$set(comment = "",echo = TRUE, fig.width = 6, fig.asp = 0.618, fig.align = "center", out.width = "70%" )
library(rmarkdown, tinytex)
library(ggplot2)
library(scales)
library(lubridate)
```

***

## Problem 1 [CO_COVID19]
Load the dataset CO_COVID19.csv

```{r}
covid <- "CO_COVID19.csv" |>
  read.csv() |>
  dplyr::as_tibble()
```

### Problem 1a
Draw a boxplot for the cumulative number of cases.
```{r}
ggplot(covid, aes(CumulativeCases)) + 
  geom_boxplot(fill = "navy") +
  labs(title = "Covid19 Cases", subtitle = "Cumulative Cases")
```

### Problem 1b

Draw a boxplot for the cumulative number of deaths.
```{r}
ggplot(covid, aes(CumulativeDeaths)) + 
  geom_boxplot(fill = "violet") +
  labs(title = "Covid19 Deaths", subtitle = "Cumulative Deaths")
```

### Problem 1c
Draw a scatterplot of the daily number of cases versus the daily number of deaths.
```{r}
ggplot(covid) +
  aes(x = DailyCases, y = DailyDeaths) +
  geom_point(shape = 23) +
  geom_smooth(method = "lm", color = "red") +
  xlab("Daily Cases") +
  ylab("Daily Deaths") +
  ggtitle("Daily Covid19 Cases vs. Deaths")
```


### Load the given dataset


```{r}
covid2 <- "CovidMonthlyData.csv" |>
  read.csv() |>
  dplyr::as_tibble()

# Clean up the date column.
covid2$Date <- covid2$Date |> my()
```


### Problem 1d
Draw a bar graph for the monthly number of cases.

```{r}
ggplot(covid2, aes(x = Date, y = Cases)) +
	geom_col(color = "black", fill = "deepskyblue") + 
  scale_y_continuous(labels = comma) +
  labs(title = "Covid19 Cases by Month", x = "Date: (Month / Year)", y = "Number of Cases") +
  theme(axis.text.x = element_text(angle = 65, vjust = 0.6))
```


### Problem 1d
Draw a bar graph for the monthly number of Deaths.

```{r}
ggplot(covid2, aes(x = Date, y = Deaths)) +
	geom_col(color = "black", fill = "deepskyblue") + 
  scale_y_continuous(labels = comma) +
  labs(title = "Covid19 Deaths per Month", x = "Date: (Month / Year)", y = "Number of Deaths") +
  theme(axis.text.x = element_text(angle = 65, vjust = 0.6))
```

### Problem 1f
Draw a pie chart for the monthly number of cases

```{r}
ggplot(covid2) +
  aes(x = "", y=Cases, fill = factor(Date)) +
  geom_col() +
  labs(fill = "Date", x=NULL, y=NULL, 
      title = "Covid19 Cases per Month") +
  coord_polar(theta = "y", start = 0) +
  scale_y_continuous(labels = comma)
```


```{r}
ggplot(covid2) +
  aes(x = "", y=Deaths, fill = factor(Date)) +
  geom_col() +
  labs(fill = "Date", x=NULL, y=NULL, 
      title = "Covid19 Deaths per Month") +
  coord_polar(theta = "y", start = 0) +
  scale_y_continuous(labels = comma)
```



\pagebreak


## Problem 2
For the mpg dataset, redo the following plots from the Week 12 Lab using the city mpg instead of the highway mpg. 

### Problem 2a
Redo the scatterplot with shape = drv, color = drv

```{r}
ggplot(mpg, aes(x = displ, y = cty, shape = drv, color = drv)) + geom_point()
```

### Problem 2b
The facet_wrap plots with 'class' and 'drv'

```{r}
ggplot(mpg, aes(x = displ, y = cty)) + geom_point(color = "purple") +
facet_wrap(~class, nrow = 2)
```

```{r}
ggplot(mpg, aes(x = displ, y = cty)) + geom_point(color = "red") + 
  facet_wrap(~drv)
```


### Problem 2c
The bar chart of make vs. average city mpg.

```{r}
make <- unique(mpg$manufacturer)
cty_mean <- tapply(mpg$cty, mpg$manufacturer, mean)
mpg2 <- data.frame(make, cty_mean)

ggplot(mpg2, aes(x = make, y = cty_mean)) +
geom_bar(stat = "identity", width = 0.5, fill = "cyan3") +
labs(title = "Bar Chart", subtitle = "Make vs. Avg City MPG",
x = "Make", y = "City Mileage MPG", caption="Source: Frequency of
Manufacturers from ’mpg’ dataset") +
theme(axis.text.x = element_text(angle = 65, vjust = 0.6))
```


### Problem 2d
The lollipop chart of make vs. average city mpg.

```{r}
ggplot(mpg2, aes(x = make, y = cty_mean)) + geom_point(size = 3) +
  geom_segment(aes(x = make, xend = make, y = 0, yend = cty_mean)) +
  labs(title = "Lollipop Chart", subtitle = "Make Vs Avg City MPG",
  x = "Make", y = "City Mileage MPG", caption="Source: Frequency of
  Manufacturers from ’mpg’ dataset") +
  theme(axis.text.x = element_text(angle = 65, vjust = 0.6))
```


\pagebreak


## Problem 3

Load the dataset

```{r}
penguins <- "penguins.csv" |>
  read.csv() |>
  dplyr::as_tibble()
```

### Problem 3a
Compute:
i) The overall average body mass
ii) The average body mass by species
iii) The average body mass by island
iv) the average body mass by sex.

```{r}
#p3i overall body mass (grams)
mean(penguins$body_mass_g)
```

```{r}
#p3ii body mass by species (grams)
tapply(penguins$body_mass_g, penguins$species, FUN = mean)
```

```{r}
# p3iii body mass by island (grams)
tapply(penguins$body_mass_g, penguins$island, FUN = mean)
```

```{r}
# p3iv body mass by sex (grams)
tapply(penguins$body_mass_g, penguins$sex, FUN = mean)
```


### Problem 3b
Draw a histogram of the body mass.

```{r}
ggplot(penguins, aes(body_mass_g)) +
  geom_histogram(color = "black", fill = "palegreen1", bins = 40) +
  labs(title = "Distribution of Penguin Body Mass",
       x = "Body Mass in Grams")
```

```{r}
ggplot(penguins, aes(body_mass_g)) +
  scale_fill_brewer(palette = "Paired") +
  geom_histogram(color = "black", aes(fill = species), binwidth = 125) +
  labs(title = "Body Mass across Penguin Species",
       x = "Body Mass in Grams")
```

```{r}
ggplot(penguins, aes(body_mass_g)) +
  scale_fill_brewer(palette = "Accent") +
  geom_histogram(aes(fill = island), binwidth = 80, color = "black") +
  labs(title = "Body Mass of Penguins across Islands",
       x = "Body Mass in Grams")
```

```{r}
ggplot(penguins, aes(body_mass_g)) +
  geom_histogram(color = "black", aes(fill = sex), binwidth = 50) +
  labs(title = "Body Mass of Penguins by Sex",
       x = "Body Mass in Grams")
```


\pagebreak


## Problem 4 [Epicycloids]
The parametric equations for the epicycloid on $[0, (R + r)\pi]$ are:

\[x = (R + r) \cos t - r \cdot \cos \left( \frac{R + r}{r} \cdot t \right)\]
\[y = (R + r) \sin t - r \cdot \sin \left( \frac{R + r}{r} \cdot t \right)\]

### Problem 4a
Modify your epicycloid function from HW 9 to include color, and replace the plot() command with the provided ggplot command. 

```{r}
Epicycloid <- function(r, R, color){
  t <- seq(from=0, to=(R + r)*2*pi, len=10000)
  
  x <- (R + r)*cos(t) - (r * cos(((R+r)/r)*t))
  # print(x)
  
  y<- (R + r)*sin(t) - (r * sin(((R+r)/r)*t)) 
  # print(y)
  
  ggplot(data.frame(t, x, y), aes(x,y)) + geom_path(color = color) +
    ggtitle(paste("Epicycloid: r =", r, ", R = ", R)) + 
    theme_void()
  }
```

```{r}
Epicycloid(r = 1, R = 4, color = "red")
```

```{r}
Epicycloid(r = 6, R = 5, color = "violetred3")
```

```{r}
Epicycloid(r = 10, R = 21, color = "red")
```

```{r}
Epicycloid(r = 32, R = 33, color = "violetred3")
```

```{r}
Epicycloid(r = 18, R = 4, color = "tomato4")
```

```{r}
Epicycloid(r = 72, R = 71, color = "sienna")
```

```{r}
Epicycloid(r = 30, R = 65, color = "purple3")
```


\pagebreak


## Problem 5 [Rose Curves]

```{r}
Rose <- function(k, color){
  t <- seq(from=0, to=(2*pi*(10/k)), by=0.001)
  
  x <- cos(k*t)*cos(t)
  y <- cos(k*t)*sin(t)
  
  
  ggplot(data.frame(t, x, y), aes(x,y)) + geom_path(color = color) +
    ggtitle(paste("Epicycloid: k =", k)) + 
    theme_void()
  }
```

```{r}
Rose(k=4, color="blue2")
```

```{r}
Rose(k=5, color="blue3")
```

```{r}
Rose(k=(3/5), color="chocolate1")
```

```{r}
Rose(k=(4/5), color="chocolate1")
```

```{r}
Rose(k=1.2, color="indianred2")
```

```{r}
Rose(k = 12, color = "purple")
```

```{r}
Rose(k=27, color = "purple")
```


\pagebreak


## Problem 6.
Make one or more 'heatmap' images.

```{r}
data <- matrix(rnorm(150), nrow=10)
heatmap(data, Colv = NA, Rowv = NA, scale = "none")
```


\pagebreak


## Problem 7

Modify the given code to make one or more variations of the artwork. 

```{r}
set.seed(345) #set the seed of R's random number generator
library(ggplot2)
library(RColorBrewer)
ngroup=32    # changes how many groups there are in the art graph
names=paste("G_",seq(1,ngroup),sep="")
DAT=data.frame()
```


```{r}
#creating dataframe 
for(i in seq(1:ngroup)){
    data=data.frame( matrix(0, ngroup , 3))
    data[,1]=cos(i) + sin(i)
    data[,2]=sample(names, nrow(data))
    data[,3]=prop.table(sample( c(rep(0,150),c(1:ngroup)) ,nrow(data)))
    DAT=rbind(DAT,data)
    }
colnames(DAT)=c("Year","Group","Value")
DAT=DAT[order( DAT$Year, DAT$Group) , ]

coul = brewer.pal(12, "Paired") 
coul = colorRampPalette(coul)(ngroup)
coul=coul[sample(c(1:length(coul)) , size=length(coul) ) ] #deciding the color for the palette
```

```{r}
ggplot(DAT, aes(x=Year, y=Value, fill=Group )) + 
    geom_area(alpha=1 , color=1 )+
    theme_bw() +
    scale_fill_manual(values = coul)+
     theme(line = element_blank(),
        text = element_blank(),
        title = element_blank(),
        legend.position="none",
        panel.border = element_blank(),
        panel.background = element_blank())
```