---
title: 'MTH-2520 Homework 6: Matrices and Arrays'
author: Brady Lamson
date: '2021-03-01'
slug: mth-2520-homework-6-matrices-and-arrays
categories: []
tags: [Homework]
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="")

library(rmarkdown, tinytex)
library(markdown)
library(dplyr)
```


***
## Problem 1. 
Create an example matrix in the three following ways:

###
Problem 1a
Using matrix()

```{r}
matrix(1:5, nrow = 5, ncol=8)
```


###
Problem 1b
Using rbind()

```{r}
matrix_1b1 <- matrix(1, nrow = 1, ncol=8, byrow=FALSE) 
matrix_1b2 <- matrix(2, nrow = 1, ncol=8, byrow=FALSE)
matrix_1b3 <- matrix(3, nrow = 1, ncol=8, byrow=FALSE)
matrix_1b4 <- matrix(4, nrow = 1, ncol=8, byrow=FALSE)
matrix_1b5 <- matrix(5, nrow = 1, ncol=8, byrow=FALSE)

rbind(matrix_1b1, matrix_1b2, matrix_1b3, matrix_1b4, matrix_1b5)
```


###
Problem 1c
Using cbind()

```{r}
matrix_1c1 <- matrix(1:5, nrow = 5, ncol=4) 
matrix_1c2 <- matrix(1:5, nrow = 5, ncol=4)

cbind(matrix_1c1, matrix_1c2)
```


***
## Problem 2.
For the matrix A in pr. 1: , b) 

### Problem 2a.
add a row of 6’s on the bottom

```{r}
matrix_2 <- matrix(1:5, nrow = 5, ncol=8)
rbind(matrix_2, rep(6, 8))
```

### Problem 2b.
then extract the middle 2 × 4 matrix

```{r}
# Before we can do this we need to know what the middle even is! This matrix is 6 rows long and 8 columns long. That means the middle 2 x 4 matrix would be in rows 3-4 and columns 3-6. We should get all 3s and 4s in our matrix.

matrix_2[row=3:4, col=3:6]
```


***
## Problem 3
Build a 10 x 10 identity matrix. Then change all the non-zero elements to 5.

```{r}
matrix_3 <- diag(10)
matrix_3 * 5
```

So, funfact for this one. I was racking my brain thinking of some way to change the values in the matrix based on a conditional (ie: where(x != 0)). My roommate walks over, looks at my screen and says "why don't you just multiply the matrix by 5". Right. Of course. The solution that's directly in the lab notes. I should probably go with that solution. 


***
## Problem 4.
Linear Algebra

### Problem 4a.
Use the random number generator rnorm() with the given seed to generate A.

```{r}
set.seed(1234)
A <- matrix(rnorm(25), nrow = 5)
A
```


### Problem 4b.
Calculate the determinant of A.

```{r}
det(A)
```


### Problem 4c.
Calculate the inverse of A.

```{r}
solve(A)
```


***
## Problem 5. 
For the given matrix compute each part separately. If you alter the matrix be sure to reset it back to its default value. 

$A = \begin{bmatrix}
1 & 1 & 1 \\
5 & 2 & 6 \\
-2 & -1 & -3
\end{bmatrix}$
```{r}
p5_vector <- c(1,1,3,5,2,6,-2,-1,-3)
p5_matrix <- matrix(p5_vector, nrow = 3, ncol = 3, byrow=TRUE) 
p5_matrix
```

### Problem 5a.
Compute $A^2$

```{r}
p5_matrix %*% p5_matrix
```


### Problem 5b. 
Compute A^2. What is R computing here?

```{r}
p5_matrix^2
```
p5_matrix^2 is doing element-wise multiplication. That is, element [1,1] is multiplied with element [1,1], element [1,2] with element [1,2] and so on. In other words, every element is multiplied with itself. This could not be more different from matrix multiplication.


### Problem 5c. 
Verify that $A^3 = 0$ where $0$ is a 3 X 3 matrix with every entry equal to 0.

```{r}
p5_matrix %*% p5_matrix %*% p5_matrix
```


### Problem 5d.
Compute A^3. What is R computing here?

```{r}
p5_matrix^3
```
Yet again we're seeing the result of element wise multiplication. Each element is simply multiplied with itself twice. 


### Problem 5e. 
Replace the 3rd column of A with the sum of the 2nd and 3rd columns. 

```{r}
columns_2_3 <- p5_matrix[,2] + p5_matrix[,3]

p5_matrix[,1] <- columns_2_3
p5_matrix
```


### Problem 5f
Compute the column means of "A". 

```{r}
cat("the column means of p5_matrix are [", colMeans(p5_matrix), "] \n")
```


### Problem 5g
Compute the square roots of the 2nd column.

```{r}
cat("The square roots of column two are [", sqrt(as.complex(p5_matrix[,2])), "] \n")
```


### Problem 5h
Subtract 5 from column 1, subtract 4 from column 2 and add 9 to column 3 of A. Use sweep().

```{r}
sweep(p5_matrix, MARGIN = 2, c(-5, -4, 9), "+") #Margin=2 in this case because we're manipulating by column not by row. 
```


***
## Problem 6
Use the outer() function to make a table (actually a matrix) of powers of integers $x^y$ where x <- 1:10 and y <- 1:9

```{r}
pr6_x <- 1:10
pr6_y <- 1:9

outer(pr6_x, pr6_y, FUN = "^")
```


*** 
## Problem 7

### Problem 7a
Create a 6 X 10 matrix of random integers from 1:10.

```{r}
set.seed(75)
p7_matrix <- matrix(sample(x = 10, size = 60, replace = TRUE), nrow = 6) 
p7_matrix
```


### Problem 7b
Find the number of entries in each row which are greater than 4. 

```{r}
# length(which(p7_matrix[1,] > 4))

greater_than_4 <- function(matrix) {
  # numrows is set to the amount of rows in a given matrix.
  numrows <- nrow(matrix)
  # We set our return vector to an empty vector that we will fill later.
  ret_vec <- c()
  # This for loop goes from 1 to the number provided by numrows.
  for (row in c(1:numrows)){
    # We filter out the amount of numbers greater than 4 per row and then count how many there are. 
    filter_var <- length(which(matrix[row,] > 4))
    # We then append that counted length to our empty vector
    ret_vec <- append(ret_vec, filter_var)
  } 
  return(ret_vec)
}

greater_than_4(p7_matrix)
```

I, for the life of me, could not figure out how to do this without a for loop. I'm positive there has to be a cleaner way to do this problem but I spent forever on it with no luck.


### Problem 7c
Find which rows contain exactly two occurrences of the number 7.

```{r}

sevens <- function(matrix) {
  
  numrows <- nrow(matrix)
  for (row in c(1:numrows)){
    # This sets seven_var to the length of a vector made up of however many times a value in a row is equal to 7.
    sevens_var <- length(which(matrix[row,] == 7))
    # This returns the row which gave us a sevens_var value of 2.
    if(sevens_var == 2){return(row)}  
  } 
}

cat("The row in p7_matrix that has two occurrences of the number 7 is row", sevens(p7_matrix), "\n")
```


### Problem 7d
Use apply() to sort each column of A in decreasing order

```{r}
apply(p7_matrix, MARGIN = 2, FUN = sort, decreasing=TRUE)
```


### Problem 7e
Use order() to sort A in decreasing order based on column 1. 

```{r}
# p7_matrix
a <- p7_matrix[order(p7_matrix[,1], decreasing=TRUE),]

a
```
This mysterious floating comma at the end being the one thing making this work correctly is absolutely blowing my mind right now. 

### Problem 7f
Use apply() and the transpose to sort each row of A in increasing order.

```{r}
mat_7f <- t(p7_matrix) 
apply(mat_7f, MARGIN = 1, FUN = sort, decreasing=FALSE)

```


### Problem 7g
Use order() to sort A in increasing order based on row 3

```{r}
# p7_matrix
b <- p7_matrix[, order(p7_matrix[3,], decreasing = FALSE)]

b
```
It took me a solid hour to figure out that I needed the comma after the [] there instead of at the end. That is, quite possibly, the silliest thing I have seen in my life (so far). The mysterious floating comma will live on in my nightmares. I fear I have not yet seen the last of its treachery. 

***
## Problem 8.
Find the sum

### Problem 8a.
\[\sum_{i=1}^{20} \sum_{j=1}^{15} (2i + 3j)\]

```{r}
# First we'll create our two elements, i and j.
var_8a_x <- c(1:20) * 2
var_8a_y <- c(1:15) * 3

sum(outer(var_8a_x, var_8a_y, FUN = "+"))
```


### Problem 8b.
\[ \sum_{i=1}^{20} \sum_{j=1}^{5} \frac{i^4}{3 + j} \]

```{r}
# We'll use a similar breakdown for this guy.
var_8b_x <- c(1:20)^4
var_8b_y <- c(1:5) + 3

sum(outer(var_8b_x, var_8b_y, FUN = "/"))

```


***
## Problem 9
Write a function that takes a matrix and returns a matrix with every odd number doubled.

```{r}
doubling_odds <- function(x){
  ifelse(x %% 2 != 0, x * 2, x)
}

pr_9a <- matrix(c(1,-3,-2,4), nrow=2)
pr_9b <- matrix(c(1:15), nrow = 3)

doubling_odds(pr_9a)
doubling_odds(pr_9b)
```


***
## Problem 10. 

### Problem 10a.
Write a function that takes a matrix and returns the smallest number in the matrix. 

```{r}

smallest_number <- function(matrix) {
  # This sets var_10a as a given matrix sorted in ascending order
  var_10a <- sort(matrix, decreasing=FALSE,)
  # This returns the first index in said matrix, which due to the above will be the smallest number.
  return(var_10a[1])
  } 
```


### Problem 10b.
Test the function on the following matrices. 

```{r}
matrix_10b1 <- matrix(c(1, -3, -2, 4), nrow = 2)
smallest_number(matrix_10b1)
smallest_number(A)
```


***
## Problem 11
Solve the system of linear equations. Verify that $det A \neq 0$. 

```{r}
matrix_11_left <- matrix(c(4, 1, 2, -3, -3, 3, -1, 4, -1, 2, 5, 1, 5, 4, 3, -1), byrow = TRUE, nrow = 4)
matrix_11_right <- matrix(c(-16, 20, -4, -10), nrow=4)

cat("det A is equal to", det(matrix_11_left, matrix_11_right), "\n")
matrix(solve(matrix_11_left, matrix_11_right), dimnames = list(c("w =", "x =", "y =", "z =")))

```

