---
title: 'MTH-2520 Homework 5: Functions and Basic Programming'
author: Brady Lamson
date: '2021-02-22'
slug: mth-2520-homework-5-functions-and-basic-programming
categories: []
tags: [Homework]
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="")
library(rmarkdown, tinytex)

```


***
## Problem 1: 
Write a function that computes the area of a circle.
```{r}
Calculate_Area <- function(r) {
# This function calculates the area of a circle
area <- pi * r^2
return(area)
}

radii <- c(5, 15, 200)

Calculate_Area(radii)
```


***
## Problem 2:
Write a function that computes the average absolute value. $\frac{\lvert x \rvert + \lvert y \rvert}{2}$.

```{r}
Average_Absolutely <- function(x, y) {
  # This function calculates the average absolute value. 
  # (|x| + |y|) / 2
  
  absolute_average <- (abs(x) + abs(y)) / 2
  return(absolute_average)
}

example_x <- c(-3:2)
example_y <- c(-3:2)

Average_Absolutely(example_x, example_y)
```


***
## Problem 3:
Write a function to count the number of odd integers in a given vector x. Test your function of the vectors: c(1, 3, 5, 6, 8, 9), seq(1, 99, 3) and seq(1, 5000, 7). 

```{r}
# Returns the amount of odd integers in a given vector
Count_Odds <- function(x) {
  odds <- length(which(x %% 2 > 0))
  return(odds)
}

ex_1 <- c(1, 3, 5, 6, 8, 9)
ex_2 <- seq(1, 99, 3)
ex_3 <- seq(1, 5000, 7)

cat("The amount of odds in the first vector are", Count_Odds(ex_1), "\n")
cat("The amount of odds in the second vector are", Count_Odds(ex_2), "\n")
cat("The amount of odds in the third vector are", Count_Odds(ex_3), "\n")
```


***
## Problem 4: 
Use ifelse or switch to return "even" or "odd" depending whether or not each element of the vector x is divisible by 2. 

```{r}
# This function finds the remainder of various integers and assigns them even or odd depending on said remainder. 
Even_or_Odd <- function(x) {
  y <- x%%2
  
  if (y == 0){
    return("even")}
  else{
    return("odd")}
}


x_1 <- c(1, 3, 5, 6, 8, 9)
x_2 <- 1:20


sapply(x_1, Even_or_Odd)
sapply(x_2, Even_or_Odd)
```


***
## Problem 5:
We wish to compute the product of 3, 4 or 5 digits, and then add 42. Write a function to complete this task. Test your function on the vectors: c(3,5,7), c(3,5,7,9) and c(5,6,7,8,9).

```{r}
sqornshellous_zeta <- function(x) {
  # This function does what it says on the tin. That is, to say, takes the product of a vector and adds the answer to life, the universe, and everything.
  
  # This function only works for vectors made up of integers, floats and logicals.
  
  product_value <- prod(x)
  final_product_value <- product_value + 42
  return(final_product_value)
}

ex_51 <- c(3,5,7)
ex_52 <- c(3,5,7,9)
ex_53 <- c(5,6,7,8,9)

sqornshellous_zeta(ex_51)
sqornshellous_zeta(ex_52)
sqornshellous_zeta(ex_53)
```


***
## Problem 6:
The sum formula for the finite geometric series $h(x,n) = \sum_{i=0}^{n} x^{i} = 1 + x + x^2 + ... + x^n, x \neq 1$ is given by: 

\[
  h(x,n) = 
  \begin{cases}
                                   \frac{1 - x^{n+1}}{1-x} & \text{for $x \neq 1$} \\
  n & \text{for $x = 1$}
  \end{cases}
\]

Write a function to compute this formula. Use your function to compute: h(0.7, 57), h(4.2, 9), h(-1.2, 81) and h(1, 342).

```{r}
piecewise_example <- function(x,n) {
  if (x == 1){return(n)}
  else{return({(1 - x^(n+1)) / (1-x)})}
}

piecewise_example(0.7, 57)
piecewise_example(4.2, 9)
piecewise_example(-1.2, 81)
piecewise_example(1, 342)
```


***
## Problem 7.
Write a function that computes $f(n) = n!$ using the formula $n! n(n-1,)(n-2) ... 2 \cdot 1$ and does not use factorial(). For a given integer $n$, the function prints out:

1) n! n> 0 
2) 1 for n=0
3) NaN for n<0
Test on n=5, 15, 0, -8

```{r}
better_factorial <- function(n){

  # Here I have if statements that return 1 if n=0 and a NaN if n<0. I put these at the start to catch them as soon as possible
  if (n == 0){return(1)}
  if (n < 0){return(NaN)}
  
  return_value <- 1
    
  # This while loop takes a value which is set to 1 and multiplies it by a variable "n" that gets progressively smaller by 1 with every loop. This loop stops once n is no longer greater than 1.
  while (n > 1) {
    return_value <- return_value * n
    n <- n-1
  }
  return(return_value)
}

cat("5! is", better_factorial(5), "\n")
cat("15! is", better_factorial(15), "\n")
cat("0! is", better_factorial(0), "\n")
cat("-8! is", better_factorial(-8), "\n")
```



***
## Problem 8
Write a function that prints out the 1st n Lucas numbers for a given n. Test your function for
n = 10, 25.

```{r}
lucas_func <- function(n){
  
  #First I need to make sure no negative values are plugged in here. 
  if (n < 1){
    return(NaN)
  }
  
  # Then we just need to setup the first two known values of this sequence. That is, 2 and 1.
  if (n == 1){
    return(2)
  }
  
  if (n == 2){
    return(1)
  }
  
  #Essentially my game plan is to have a and b add with each other to create a variable "c". After doing that we reassign some values. The key here is the counter. It goes up by 1 at the end of the loop and when it is no longer less than n the loop stops. 
  a <- 2
  b <- 1
  counter <- 2
  results_list <- c(2, 1) # We create a small vector here with our two given values.
  
  while (counter < n) {
    c <- a + b
    a <- b
    b <- c
    counter <- counter + 1
    results_list <- append(results_list, c) #This makes that small little vector useful! 
  }
return(results_list)
}

test_val <- c(10, 25)
sapply(test_val, lucas_func)
```


***
## Problem 9

```{r}
tribo_func <- function(n){
  
  #Essentially my gameplan here is the exact same as problem 8, I just add in another variable. 
  
  if (n < 1){
    return(NaN)
  }
  
  if (n == 1){
    return(0)
  }
  
  if (n == 2){
    return(0)
  }
  
  if (n == 3){
    return(1)
  }
  
  a <- 0
  b <- 0
  c <- 1
  counter <- 3
  results_list <- c(0, 0, 1)
  
  while (counter < n) {
    d <- a + b + c
    a <- b
    b <- c
    c <- d
    counter <- counter + 1
    results_list <- append(results_list, d)
  }
return(results_list)
}

test_val2 <- c(10, 25)
sapply(test_val2, tribo_func)
```


***
# Problem 10
Using the code for the Fibonacci sequence given in class, generate a vector containing
every 3rd term of the 1st 30 terms. Then verify using R that is vector contains only even
numbers.
Theorem: Every third Fibonacci number is even.

```{r}
Fibonacci <- function(n){
  
fib <- numeric(n)
# Create a vector of length n

fib[1] <- fib[2] <- 1
# Store F 1 and F 2 in the vector

for (i in 3:n) fib[i] = fib[i-1] + fib[i-2]

cat("The first", n, "Fibonacci numbers are", "\n")
print(fib)}

# Here I set fib_30 to the vector created when 30 is put into the fibonacci function.
fib_30 <- Fibonacci(30)
# From here I filter it out using a sequence. The sequence starts at the third position and moves up by 3. 
fib_thirds <- fib_30[seq(3, length(fib_30), 3)]

cat("Every third value in the Fibonacci sequence up to the thirtieth are", fib_thirds, "\n")

# To check if all the values here are even we can actually use our old Even_or_Odd function!
sapply(fib_thirds, Even_or_Odd)
```


***
# Problem 11
Using the code for the Fibonacci sequence given in class, generate a vector containing
every 4th term of the 1st 30 terms. Then verify using R that every number in this vector is divisible by three.
Theorem: Every fourth Fibonacci number is divisible by three.

```{r}
Fibonacci <- function(n){
  
fib <- numeric(n)
# Create a vector of length n

fib[1] <- fib[2] <- 1
# Store F 1 and F 2 in the vector

for (i in 3:n) fib[i] = fib[i-1] + fib[i-2]

cat("The first", n, "Fibonacci numbers are", "\n")
print(fib)}


# Here I set fib_30 to the vector created when 30 is put into the fibonacci function.
fib_30 <- Fibonacci(30)

# From here I filter it out using a sequence. The sequence starts at the fourth position and moves up by 4. 
fib_fourth <- fib_30[seq(4, length(fib_30), 4)]

cat("Every fourth value in the Fibonacci sequence up to the thirtieth are", fib_fourth, "\n")


# To see if all of these are divisible by three we just slightly tweak our even or odd function!

divisible_by_three <- function(x) {
  y <- x%%3
  
  if (y == 0){
    return("TRUE")}
  else{
    return("FALSE")}
}

sapply(fib_fourth, divisible_by_three)
```



***
## Problem 12
Generate a sequence of ratios with the Fibonacci numbers. List the 1st 30 terms of this sequence and verify the sequence appears to converge to 1.618034.

```{r}

fibonacci_ratio <- function(n){
  
  fib_vector <- Fibonacci(n+1) #I add the +1 here to prevent things from getting wacky when we reach the end of the final loop.
  ratio_vector <- c()
  
  for (i in 0:n) ratio_vector[i] = fib_vector[i+1] / fib_vector[i] #The n on this line could be replaced with n-1 if you remove the +1 from the n a few lines up. 
  
  return(ratio_vector)
}

fibonacci_ratio(30)
```